Apprentissage par imitation dans un cadre \emph{batch}, \emph{off-policy} et sans modèle

### 2011-06-06 main.tex
Corrections suggérées par Yann

\maketitle
\begin{abstract}
 Ce papier traite le problème de l'apprentissage par imitation, c'est à dire la résolution du problème du contrôle optimal à partir de données tirées de démonstrations d'un expert. L'apprentissage par renforcement inverse (IRL) propose un cadre efficace pour résoudre ce problème. En se basant sur l'hypothèse que l'expert maximise une fonction de valeur,  l'IRL essaie d'apprendre la récompense qui définit cette dernière à partir de trajectoires d'exemple. Beaucoup d'algorithmes d'IRL font l'hypothèse de l'existence d'unn approximateur linéaire pour la fonction de récompense et calculent l'attribut moyen (le cumul moyen pondéré des fonctions de base, relatives à la paramétrisation linéaire supposée de la récompense, évaluées en les états d'une trajectoire associée à une certaine politique)  via une estimation de Monte-Carlo. Cela implique d'avoir accès à des trajectoires complètes de l'expert ainsi qu'à au moins un modèle génératif pour tester les politiques intermédiaires. Dans ce papier nous introduisons une méthode de différences temporelles, LSTD-$\mu$, pour calculer cet attribut moyen. Cela permet d'étendre l'apprentissage par imitation aux cas \emph{batch} et \emph{off-policy}.
\end{abstract}
\section{Introduction}

@@

There roughly exists three families of IRL algorithms:
feature-expectation-based~\citep{abbeel2004apprenticeship,syed2008apprenticeship,syed2008game,ziebart2008maximum},
margin-maximization-based~\citep{ratliff2006maximum,ratliff2007imitation,ratliff2007boosting,kolter2008hierarchical}
and approaches based on the parameterization of the policy by the
reward function~\citep{ramachandran2007bayesian,neu2007apprenticeship}. The first family
assumes a linearly parameterized reward function. This naturally
leads to a linearly parameterized value function, the associated
feature vector being the so-called feature expectation (defined as
the cumulative discounted reward's feature vector under a given
policy, see Section~\ref{sec:back} for a formal definition).
These approaches learn a reward function such that the feature
expectation of the optimal policy (according to the learnt reward
function) is close to the feature expectation of the expert policy.
This is a sufficient condition to have close value functions (for any parameterized reward function, and therefore particularly the optimized one). The
second family expresses IRL as a constrained optimization problem in
which expert's examples have higher expected cumulative reward than all other
policies by a certain margin. Moreover, suboptimality of the expert
can be considered through the introduction of slack variables. The
last family parameterizes policies with a reward function. Assuming
that the expert acts according to a Gibbs policy (respectively to
the optimal value function related to the reward function which is
optimized), it is possible to estimate the likelihood of a set of
state-action pairs provided by the expert. The algorithms differs in
the way this likelihood is maximized.

This paper focuses on the first family of algorithms, and more
precisely on the seminal work of~\citet{abbeel2004apprenticeship}. All of them
rely on the computation of the feature expectation (which depends on
policies but not on rewards) of (i) the expert and (ii) some
intermediate policies. The expert's feature expectation is computed
using a simple Monte Carlo approach (which requires full
trajectories of the expert). Other feature expectations are either
computed exactly (which requires knowing analytically the dynamics
of the system) or with a Monte Carlo approach (which requires simulating the system). The contribution of this paper is LSTD-$\mu$,
a new temporal-difference-based algorithm for estimating these feature
expectations. It relaxes the preceding assumptions: transitions
of the expert are sufficient (rather than full trajectories) and nor
the model neither a simulator are necessary to compute intermediate
feature expectations. This paper focuses on the algorithm introduced
in~\citep{abbeel2004apprenticeship}, but we are confident that the proposed
approach can be extended to other algorithms based on feature
expectation computation~\citep{syed2008apprenticeship,syed2008game,ziebart2008maximum}, or even \citep{ratliff2006maximum,ratliff2007imitation,ratliff2007boosting}.

@@

\section{Background}
\label{sec:back}
A sequential decision problem is often framed as a Markov Decision
Process (MDP)~\citep{puterman1994markov}. An MDP is a tuple
$\{S,A,P,R,\gamma\}$ with $S$ being the state space, $A$ the action
space, $P\in\mathcal{P}(S)^{S\times A}$ the set of Markovian
transition probabilities, $R\in\mathbb{R}^S$ the reward function
(assumed to be absolutely bounded by 1) and $\gamma\in[0,1[$ a
discounting factor. A policy $\pi\in A^S$ maps states to action. The
quality of a policy is quantified by the associated value function
$V^\pi$, which associates to each state the expected and discounted
cumulative reward:
\begin{equation}
  V^\pi(s) = E[\sum_{t=0}^\infty \gamma^t R(s_t)|s_0=s, \pi]
\end{equation}
Dynamic programming aims at finding the optimal policy $\pi^*$, that
is one of the policies associated to the optimal value function,
$V^* = \argmax_\pi V^\pi$, which maximizes the value for each state.
If the model (that is transition probabilities and the reward
function) is unknown, learning the optimal control through
interactions is addressed by reinforcement learning.

For inverse reinforcement learning, the problem is reversed. It is
assumed that an expert acts according to an optimal policy $\pi_E$,
this policy being optimal according to some unknown reward function
$R^*$. The goal of inverse reinforcement learning is to learn this
reward function from sampled trajectories of the expert. This is a
difficult and ill-posed problem~\citep{ng2000algorithms}. Apprenticeship
learning through IRL, which is the problem at hand in this paper,
has a somewhat weaker objective: it aims at learning a policy
$\tilde{\pi}$ which is (approximately) as good as the expert policy
$\pi_E$ under the unknown reward function $R^*$, for a known initial
state $s_0$ (this condition can be weakened by assuming a
distribution over initial states ; this is not done here for clarity
of exposition). Now, the approach proposed in~\citep{abbeel2004apprenticeship} is
presented.

We assume that the true reward funnotiction belongs to some hypothesis
space $\mathcal{H}_\phi = \{\theta^T \phi(s),
\theta\in\mathbb{R}^p\}$, of which we assume the basis functions to
be bounded by 1: $|\phi_i(s)|\leq 1, \forall s\in S, 1\leq i \leq
p$. Therefore, there exists a parameter vector $\theta^*$ such that:
\begin{equation}
  R^*(s) = (\theta^*)^T \phi(s)
\end{equation}
In order to ensure that rewards are bounded, we impose that
$\|\theta\|_2\leq 1$. For any reward function belonging to
$\mathcal{H}_\phi$ and for any policy $\pi$, the related value
function $V^\pi(s)$ can be expressed as follows:
\begin{align}
  V^\pi(s) &= E[\sum_{t=0}^\infty \gamma^t \theta^T \phi(s_t)|s_0=s, \pi]
  \\
  &= \theta^T  E[\sum_{t=0}^\infty \gamma^t \phi(s_t)|s_0=s, \pi]
\end{align}
Therefore, the value function is also linearly parameterized, with
the same weights and with basis functions being grouped into the
so-called \emph{feature expectation} $\mu^\pi$:
\begin{equation}
  \mu^\pi(s) = E[\sum_{t=0}^\infty \gamma^t \phi(s_t)|s_0=s, \pi]
\end{equation}
Recall that the problem is to find a policy whose performance is
close to that of the expert's for the starting state $s_0$, on the
unknown reward function $R^*$. In order to achieve this goal, it is
proposed in~\citep{abbeel2004apprenticeship} to find a policy $\tilde{\pi}$ such
that $\|\mu^{\pi_E}(s_0)-\mu^{\tilde{\pi}}(s_0)\|_2\leq \epsilon$
for some (small) $\epsilon>0$. Actually, this ensures that the value
of the expert's policy and the value of the estimated policy (for
the starting state $s_0$) are close for \emph{any} reward function
of $\mathcal{H}_\phi$:
\begin{align}
  |V^{\pi_E}(s_0) - V^{\tilde{\pi}}(s_0)| &=
  |\theta^T(\mu^{\pi_E}(s_0)-\mu^{\tilde{\pi}}(s_0))|
  \label{eqn:vs0}
  \\
  &\leq \|\mu^{\pi_E}(s_0)-\mu^{\tilde{\pi}}(s_0)\|_2
\end{align}
This last equation uses the Cauchy-Schwartz inequality and the
assumption that $\|\theta\|_2\leq 1$. Therefore, the approach
described here does not ensure to retrieve the true reward function
$R^*$, but to act as well as the expert under this reward function
(and actually under any reward function, notably the unknown optimal
one).

Let us now describe the algorithm proposed in~\citep{abbeel2004apprenticeship} to
achieve this goal:
\begin{enumerate}
  \item Starts with some initial policy $\pi^{(0)}$ and compute
  $\mu^{\pi^{(0)}}(s_0)$. Set $j=1$;
  %
  \item Compute $t^{(j)} = \max_{\theta: \|\theta\|_2\leq 1}
  \min_{k\in\{0,j-1\}}\theta^T(\mu^{\pi_E}(s_0) -
  \mu^{\pi^{(k)}}(s_0))$ and let $\theta^{(j)}$ be the value
  attaining this maximum. At this step, one searches for the reward
  function which maximizes the distance between the value of the
  expert at $s_0$ and the value of \emph{any} policy computed so far
  (still at $s_0$). This optimization problem can be solved using a
  quadratic programming approach or a projection
  algorithm~\citep{abbeel2004apprenticeship};
  %
  \item if $t^{(j)}\leq \epsilon$, terminate. The algorithm outputs a
  set of policies $\{\pi^{(0)}, \dots, \pi^{(j-1)}\}$ among which
  the user chooses manually or automatically the closest to the
  expert (see~\citep{abbeel2004apprenticeship} for details on how to choose this
  policy\footnote{A simple solution, not mentioned in~\citep{abbeel2004apprenticeship},
  is to choose the policy of which the feature expectation is the
  closest to the one of the expert.}).
  Notice that the last policy is not necessary the best (as illustrated in Section \ref{sec:exp});
  %
  \item solve the MDP with the reward function $R^{(j)}(s) =
  (\theta^{(j)})^T\phi(s)$ and denote $\pi^{(j)}$ the associated
  optimal policy. Compute $\mu^{\pi^{(j)}}(s_0)$;
  %
  \item set $j\leftarrow j+1$ and go back to step 2.
\end{enumerate}
There remains three problems: computing the feature expectation of
the expert, solving the MDP and computing feature expectations of
intermediate policies.

@@

\section{LSTD-$\mu$}
\label{sec:lstdmu}
Let us write the definition of the $i^\text{th}$ component of a
feature expectation $\mu^\pi(s)$ for some policy $\pi$:
\begin{equation}
  \mu_i^\pi(s) = E[\sum_{t=0}^\infty \gamma^t \phi_i(s_t)|s_0=s,\pi]
  \label{eqn:phi}
\end{equation}
This is exactly the definition of the value function of the policy
$\pi$ for the MDP considered with the $i^\text{th}$ basis function
$\phi_i(s)$ as the reward function. There exists many algorithms to
estimate a value function, any of them can be used to estimate
$\mu_i^\pi$. Based on this remark, we propose to use specifically
the least-squares temporal difference (LSTD)
algorithm~\citep{bradtke1996linear} to estimate each component of the
feature expectation (as each of these components can be understood
as a value function related to a specific and known reward
function).

@@

There remains to compute the feature expectations of intermediate
policies, which should be done in an off-policy manner (that is
without explicitly sampling trajectories according to the policy of
interest). To do so, still interpreting each component of the
feature expectation as a value function, we introduce a state-action
feature expectation defined as follows (much as the classic
$Q$-function extends the value function):
\begin{equation}
  \mu^\pi(s,a) = E[\sum_{t=0}^\infty \gamma^t
  \phi(s_t)|s_0=s,a_0=a,\pi]
\end{equation}
Compared to the classic feature expectation, this definition adds a
degree of freedom on the first action to be chosen before following
the policy $\pi$. With a slightly different definition of the
related hypothesis space, each component of this feature expectation
can still be estimated using the LSTD algorithm (namely using the
LSTD-Q algorithm~\citep{lagoudakis2003least}). The clear advantage of
introducing the state-action feature expectation is that this
additional degree of freedom allows off-policy learning. Assuming
that $\mathcal{H}_\psi$ defines state-action feature vectors and
that a set of transitions $\{(s_t,a_t,r_t,s_{t+1})_{1\leq t \leq
n}\}$ sampled according to some behaviorial policy $\pi_0$ is
available, the parameter vector $\xi_i^*$ related to the
$i^\text{th}$ component of the state-action feature expectation
$\mu^\pi(s)$ (policy $\pi$ being different from policy $\pi_0$) is
given by:
\begin{equation}
  \xi_i^* = \left(\sum_{t=1}^n
  \psi(s_t,a_t)(\psi(s_t,a_t)-\gamma\psi(s'_{t},\pi(s'_t)))^T\right)^{-1}
  \sum_{t=1}^n \psi(s_t,a_t) \phi_i(s_t)
\end{equation}
Extending LSTD-$\mu$ to state-action LSTD-$\mu$ is done in the same
manner as LSTD is extended to LSTD-Q~\citep{lagoudakis2003least},
technical details are not provided here for the clarity of
exposition.

Given the (state-action) LSTD-$\mu$ algorithm, the apprenticeship
learning algorithm of~\citep{abbeel2004apprenticeship} (see
Section~\ref{sec:back}) can be easily extended to a batch and
off-policy setting. The solely available data is a set of
transitions sampled according to the expert policy  (and possibly a set of sub-optimal but good trajectories). The
corresponding feature expectation for the starting state $s_0$ is
estimated with the LSTD-$\mu$ algorithm. At step~4 of this
algorithm, the MDP is (approximately) solved using
LSPI~\citep{lagoudakis2003least} (an approximate policy iteration
algorithm using LSTD-Q as the off-policy $Q$-function estimator).
The corresponding feature expectation at state $s_0$ is estimated
using the proposed state-action LSTD-$\mu$.

Before presenting some experimental results, let us stress that
LSTD-$\mu$ is simply the LSTD algorithm applied to a specific reward
function. Although quite clear, the idea of using a temporal
difference algorithm to estimate the feature expectation is new, as
far as we know. A clear advantage of the proposed approach is that
any theoretical result holding for LSTD also holds for LSTD-$\mu$,
such as convergence~\citep{nedic2003least} or
finite-sample~\citep{lazaric2010finiteLSTD} analysis for example, also, the algorithm learns from transitions and not trajectories.

\section{Experimental benchmark}
\label{sec:exp}
This section provides some preliminary experimental results on a simple maze benchmark. Subsection \ref{ssec:exp} details the protocol and the results while subsection \ref{ssec:quality} inspects the meaning of the different quality criteria. 
\subsection{Experiment description and results}
\label{ssec:exp}
The experimental benchmark chosen here is one of those proposed in \citep{ng2000algorithms}, a 5x5 grid world. The agent is in one of the cell of the grid (whose coordinates is the state) and can choose at each step one of the four compass directions (the action). With probability 0.9, the agent moves in the intended direction. With probability 0.1, the direction is randomly drawn (and thus have probability 0.25 to match the original direction). The reward optimized by the expert is 0 everywhere except in the upper-right cell, where it is 1. For every policy, an episode ends when the upper right cell is attained, or after 20 steps. At the start of each episode, the agent is in the lower-left corner of the grid (the opposite of where the reward is).\\

Both the state and action spaces are finite and of small cardinality. Hence, the chosen feature functions $\phi$ and $\psi$ (see equations (\ref{eqn:phi}) and (\ref{eqn:psi}) for a remainder of their meaning) are the typical features of a tabular representation :0 everywhere except the component corresponding to the current state (-action pair).\\

We want to compare the efficiency of both versions of the algorithm with repect to the number of samples available from the expert, as these samples usually are the bottleneck. Indeed as they are quite costly (in both means and human time) to generate they are often not in abundance hence the critical need for an algorithm to be expert-sample efficient. Having a simulator at one's disposal can also be difficult. For the simple problem we use this is however not an issue. The discussion about the choice of the performance metric has its own dedicated subsection (subsection~\ref{ssec:quality}). We use here the $||\mu^{\pi^{(j)}}(s_0) - \mu^{\pi_E}(s_0)||_2$ error term. Fig.~\ref{fig:E} shows, for some numbers of samples from the expert, the value of $||\mu^{\pi^{(j)}}(s_0)-\mu^{\pi_E}(s_0)||_2$ after the algorithm converged or attained the maximum number of iterations (we used 40). The best policy is found by LSTD variant after one iteration only\footnote{Precise reasons for why it happens are not clear now, but certainly have something to do with the fact that all the estimations are made along the same distribution of samples.} whereas in the MC variant, convergence happens after at least ten iterations. The best policy is not always the last, and although it experimentally always have been with the LSTD variant, there is absolutely no way to tell whether this is a feature. The score of the best policy (not the last) according to the $||\mu^{\pi^{(j)}}(s_0) - \mu^{\pi_E}(s_0)||_2$ error term is plotted here. We can see that although the LSTD variant is not as good as the MC variant when very few samples are available, both algorithm quickly converge to almost the same value ; our version converges to a slightly lower error value. The fact that our variant can work in a batch, off-policy and model-free setting should make it suitable to a range of task inaccessible to the MC variant, the requirement of a simulator being often constraining.
\subsubsection{Inverted Pendulum}
\begin{figure}
\centering
\resizebox{\columnwidth}{!}{\input{../Code/InvertedPendulum/threshold}}
\caption{Number of balancing steps for the policies found by both variants with respect to the number of samples from a sub-optimal policy given to LSPI. Note the absence of middle ground in the quality.}
\label{fig:threshold}
\end{figure}
\begin{figure}
\centering
\resizebox{\columnwidth}{!}{\input{../Code/InvertedPendulum/threshold_EB}}
\caption{Same plot as Fig.~\ref{fig:threshold}, with mean and standard deviation over 100 runs. The big standard deviation stems from the absence of middle ground in the quality of the policies. The increasing mean w.r.t. the abscissa means that the proportion of good policies increases with the number of samples from a sub-optimal policy given to LSPI.}
\label{fig:threshold_EB}
\end{figure}
Another classic toy problem is the \emph{inverted pendulum}, used for example in \citep{lagoudakis2003least} from where we drew the exact configuration (mass of the pendulum, length of the arm, definition of the feature space, etc.). In this setting, the machine must imitate an expert who maintains the unstable equilibrium of an inverted pendulum allowed to move along only one axis. A balancing step is considered successful (reward $0$) when the angle is less than $\pi\over 2$. It is considered a failure (reward $-1$) when the pendulum has fallen (i.e. the angle is more than $\pi \over 2$). A run is stopped after a failure or after 3000 successful balancing steps.\\

>>>>

\maketitle
\begin{abstract}
 Ce papier traite le problème de l'apprentissage par imitation, c'est à dire la résolution du problème du contrôle optimal à partir de données tirées de démonstrations d'un expert. L'apprentissage par renforcement inverse (IRL) propose un cadre efficace pour résoudre ce problème. En se basant sur l'hypothèse que l'expert maximise une fonction de valeur,  l'IRL essaie d'apprendre la récompense qui définit cette dernière à partir de trajectoires d'exemple. Beaucoup d'algorithmes d'IRL font l'hypothèse de l'existence d'un approximateur linéaire pour la fonction de récompense et calculent l'attribut moyen (le cumul moyen pondéré des fonctions de base, relatives à la paramétrisation linéaire supposée de la récompense, évaluées en les états d'une trajectoire associée à une certaine politique)  via une estimation de Monte-Carlo. Cela implique d'avoir accès à des trajectoires complètes de l'expert ainsi qu'à au moins un modèle génératif pour tester les politiques intermédiaires. Dans ce papier nous introduisons une méthode de différences temporelles, LSTD-$\mu$, pour calculer cet attribut moyen. Cela permet d'étendre l'apprentissage par imitation aux cas \emph{batch} et \emph{off-policy}.
\end{abstract}
\section{Introduction}

@@

There roughly exists three families of IRL algorithms:
feature-expectation-based~\citep{abbeel2004apprenticeship,syed2008apprenticeship,syed2008game,ziebart2008maximum},
margin-maximization-based~\citep{ratliff2006maximum,ratliff2007imitation,ratliff2007boosting,kolter2008hierarchical}
and approaches based on the parameterization of the policy by the
reward function~\citep{ramachandran2007bayesian,neu2007apprenticeship}. The first family
assumes a linearly parameterized reward function. This naturally
leads to a linearly parameterized value function, the associated
feature vector being the so-called feature expectation (see Section~\ref{sec:back} for a formal definition).
These approaches learn a reward function such that the feature
expectation of the optimal policy (according to the learnt reward
function) is close to the feature expectation of the expert policy.
This is a sufficient condition to have close value functions (for any parameterized reward function, and therefore particularly the optimized one). The
second family expresses IRL as a constrained optimization problem in
which expert's examples have higher expected cumulative reward than all other
policies by a certain margin. Moreover, suboptimality of the expert
can be considered through the introduction of slack variables. The
last family parameterizes policies with a reward function. Assuming
that the expert acts according to a Gibbs policy (respectively to
the optimal value function related to the reward function which is
optimized), it is possible to estimate the likelihood of a set of
state-action pairs provided by the expert. The algorithms differ in
the way this likelihood is maximized.

This paper focuses on the first family of algorithms, and more
precisely on the seminal work of~\citet{abbeel2004apprenticeship}. All of them
rely on the computation of the feature expectation (which depends on
policies but not on rewards) of (i) the expert and (ii) some
intermediate policies. The expert's feature expectation is computed
using a simple Monte Carlo approach (which requires full
trajectories of the expert). Other feature expectations are either
computed exactly (which requires knowing analytically the dynamics
of the system) or with a Monte Carlo approach (which requires simulating the system). The contribution of this paper is LSTD-$\mu$,
a new temporal-difference-based algorithm for estimating these feature
expectations. It relaxes the preceding assumptions: transitions
of the expert are sufficient (rather than full trajectories) and nor
the model neither a simulator are necessary to compute intermediate
feature expectations. This paper focuses on the algorithm introduced
in~\citep{abbeel2004apprenticeship}, but the proposed
approach can be used in other algorithms based on feature
expectation computation.

@@

\section{Background}
\label{sec:back}
A sequential decision problem is often framed as a Markov Decision
Process (MDP)~\citep{puterman1994markov}, a tuple
$\{S,A,P,R,\gamma\}$ with $S$ being the state space, $A$ the action
space, $P\in\mathcal{P}(S)^{S\times A}$ the set of Markovian
transition probabilities, $R\in\mathbb{R}^S$ the reward function
(assumed to be absolutely bounded by 1) and $\gamma\in[0,1[$ a
discounting factor. A policy $\pi\in A^S$ maps states to actions. The
quality of a policy is quantified by the associated value function
$V^\pi$, which associates to each state the expected and discounted
cumulative reward:
\begin{equation}
  V^\pi(s) = E[\sum_{t=0}^\infty \gamma^t R(s_t)|s_0=s, \pi]
\end{equation}
Dynamic programming aims at finding the optimal policy $\pi^*$, that
is one of the policies associated with the optimal value function,
$V^* = \argmax_\pi V^\pi$, which maximizes the value for each state.
If the model (that is transition probabilities and the reward
function) is unknown, learning the optimal control through
interactions is addressed by RL.

For IRL, the problem is reversed. It is
assumed that an expert acts according to an optimal policy $\pi_E$,
this policy being optimal according to some unknown reward function
$R^*$. The goal is to learn this
reward function from sampled trajectories of the expert. This is a
difficult and ill-posed problem~\citep{ng2000algorithms}. Apprenticeship
learning through IRL, which is the problem at hand in this paper,
has a somewhat weaker objective: it aims at learning a policy
$\tilde{\pi}$ which is (approximately) as good as the expert policy
$\pi_E$ under the unknown reward function $R^*$, for a known initial
state $s_0$ (this condition can be easily weakened by assuming a
distribution over initial states). Now, the approach proposed in~\citep{abbeel2004apprenticeship} is
presented.

We assume that the true reward function belongs to some hypothesis
space $\mathcal{H}_\phi = \{\theta^T \phi(s),
\theta\in\mathbb{R}^p\}$, of which we assume the basis functions to
be bounded by 1: $|\phi_i(s)|\leq 1, \forall s\in S, 1\leq i \leq
p$. Therefore, there exists a parameter vector $\theta^*$ such that:
\begin{equation}
  R^*(s) = (\theta^*)^T \phi(s)
\end{equation}
In order to ensure that rewards are bounded, we impose that
$\|\theta\|_2\leq 1$. For any reward function belonging to
$\mathcal{H}_\phi$ and for any policy $\pi$, the related value
function $V^\pi(s)$ can be expressed as follows:
\begin{align}
  V^\pi(s) &= E[\sum_{t=0}^\infty \gamma^t \theta^T \phi(s_t)|s_0=s, \pi]
  \\
  &= \theta^T  E[\sum_{t=0}^\infty \gamma^t \phi(s_t)|s_0=s, \pi]
\end{align}
Therefore, the value function is also linearly parameterized, with
the same weights and with basis functions being grouped into the
so-called \emph{feature expectation} $\mu^\pi$:
\begin{equation}
  \mu^\pi(s) = E[\sum_{t=0}^\infty \gamma^t \phi(s_t)|s_0=s, \pi]
\end{equation}
Recall that the problem is to find a policy whose performance is
close to that of the expert's for the starting state $s_0$, on the
unknown reward function $R^*$. In order to achieve this goal, it is
proposed in~\citep{abbeel2004apprenticeship} to find a policy $\tilde{\pi}$ such
that $\|\mu^{\pi_E}(s_0)-\mu^{\tilde{\pi}}(s_0)\|_2\leq \epsilon$
for some (small) $\epsilon>0$. Actually, this ensures that the value
of the expert's policy and the value of the estimated policy (for
the starting state $s_0$) are close for \emph{any} reward function
of $\mathcal{H}_\phi$:
\begin{align}
  |V^{\pi_E}(s_0) - V^{\tilde{\pi}}(s_0)| &=
  |\theta^T(\mu^{\pi_E}(s_0)-\mu^{\tilde{\pi}}(s_0))|
  \label{eqn:vs0}
  \\
  &\leq \|\mu^{\pi_E}(s_0)-\mu^{\tilde{\pi}}(s_0)\|_2
\end{align}
This equation uses the Cauchy-Schwarz inequality and the
assumption that $\|\theta\|_2\leq 1$. Therefore, the approach
described here does not ensure to retrieve the true reward function
$R^*$, but to act as well as the expert under this reward function
(and actually under any reward function).

Let us now describe the algorithm proposed in~\citep{abbeel2004apprenticeship} to
achieve this goal:
\begin{enumerate}
  \item Starts with some initial policy $\pi^{(0)}$ and compute
  $\mu^{\pi^{(0)}}(s_0)$. Set $j=1$;
  %
  \item Compute $t^{(j)} = \max_{\theta: \|\theta\|_2\leq 1}
  \min_{k\in\{0,j-1\}}\theta^T(\mu^{\pi_E}(s_0) -
  \mu^{\pi^{(k)}}(s_0))$ and let $\theta^{(j)}$ be the value
  attaining this maximum. At this step, one searches for the reward
  function which maximizes the distance between the value of the
  expert at $s_0$ and the value of \emph{any} policy computed so far
  (still at $s_0$). This optimization problem can be solved using a
  quadratic programming approach or a projection
  algorithm~\citep{abbeel2004apprenticeship};
  %
  \item if $t^{(j)}\leq \epsilon$, terminate. The algorithm outputs a
  set of policies $\{\pi^{(0)}, \dots, \pi^{(j-1)}\}$ among which
  the user chooses manually or automatically the closest to the
  expert (see~\citep{abbeel2004apprenticeship} for details on how to choose this
  policy).
  Notice that the last policy is not necessarily the best (as illustrated in Section~\ref{sec:exp});
  %
  \item solve the MDP with the reward function $R^{(j)}(s) =
  (\theta^{(j)})^T\phi(s)$ and denote $\pi^{(j)}$ the associated
  optimal policy. Compute $\mu^{\pi^{(j)}}(s_0)$;
  %
  \item set $j\leftarrow j+1$ and go back to step 2.
\end{enumerate}
There remain three problems: computing the feature expectation of
the expert, solving the MDP and computing feature expectations of
intermediate policies.

@@

\section{LSTD-$\mu$}
\label{sec:lstdmu}
Let us write the definition of the $i^\text{th}$ component of a
feature expectation $\mu^\pi(s)$ for some policy $\pi$:
\begin{equation}
  \mu_i^\pi(s) = E[\sum_{t=0}^\infty \gamma^t \phi_i(s_t)|s_0=s,\pi]
  \label{eqn:phi}
\end{equation}
This is exactly the definition of the value function of the policy
$\pi$ for the MDP considered with the $i^\text{th}$ basis function
$\phi_i(s)$ as the reward function. There exist many algorithms to
estimate a value function, any of them can be used to estimate
$\mu_i^\pi$. Based on this remark, we propose to use specifically
the least-squares temporal difference (LSTD)
algorithm~\citep{bradtke1996linear} to estimate each component of the
feature expectation (as each of these components can be understood
as a value function related to a specific and known reward
function).

@@

There remains to compute the feature expectations of intermediate
policies, which should be done in an off-policy manner (that is
without explicitly sampling trajectories according to the policy of
interest). To do so, still interpreting each component of the
feature expectation as a value function, we introduce a state-action
feature expectation defined as follows (much as the classic
$Q$-function extends the value function):
\begin{equation}
  \mu^\pi(s,a) = E[\sum_{t=0}^\infty \gamma^t
  \phi(s_t)|s_0=s,a_0=a,\pi]
\end{equation}
Compared to the classic feature expectation, this definition adds a
degree of freedom on the first action to be chosen before following
the policy $\pi$. With a slightly different definition of the
related hypothesis space, each component of this feature expectation
can still be estimated using the LSTD algorithm (namely using the
LSTD-Q algorithm~\citep{lagoudakis2003least}). The clear advantage of
introducing the state-action feature expectation is that this
additional degree of freedom allows off-policy learning.
Extending LSTD-$\mu$ to state-action LSTD-$\mu$ is done in the same
manner as LSTD is extended to LSTD-Q~\citep{lagoudakis2003least},
technical details are not provided here for the clarity of
exposition.

Given the (state-action) LSTD-$\mu$ algorithm, the apprenticeship
learning algorithm presented in Section~\ref{sec:back} can be easily extended to a batch and
off-policy setting. The solely available data is a set of
transitions sampled according to the expert policy  (and possibly a set of sub-optimal trajectories). The
corresponding feature expectation for the starting state $s_0$ is
estimated with the LSTD-$\mu$ algorithm. At step~4 of this
algorithm, the MDP is (approximately) solved using
LSPI~\citep{lagoudakis2003least} (an approximate policy iteration
algorithm using LSTD-Q as the off-policy $Q$-function estimator).
The corresponding feature expectation at state $s_0$ is estimated
using the proposed state-action LSTD-$\mu$.

Before presenting some experimental results, let us stress that
LSTD-$\mu$ is simply the LSTD algorithm applied to a specific reward
function. Although quite clear, the idea of using a temporal
difference algorithm to estimate the feature expectation is new, as
far as we know. A clear advantage of the proposed approach is that
any theoretical result holding for LSTD also holds for LSTD-$\mu$,
such as convergence~\citep{nedic2003least} or
finite-sample~\citep{lazaric2010finiteLSTD} analysis for example.
\section{Experimental benchmark}
\label{sec:exp}
This section provides experimental results on two complementary problems. Subsection~\ref{ssec:exp} details the protocol and the results while subsection~\ref{ssec:quality} inspects the meaning of the different quality criteria. 
\subsection{Experiment description and results}
\label{ssec:exp}
\subsubsection{GridWorld}
\begin{figure}
\centering
\resizebox{\columnwidth}{!}{\input{../Code/GridWorld/both_error_EB}}
\caption{$||\mu^{\pi}(s_0) - \mu^{\pi_E}(s_0)||_2$ with respect to the number of samples available from the expert. The error bars represent the standard deviation over 100 runs.}
\label{fig:E}
\end{figure}
The first experimental benchmark chosen here is one of those proposed in \citep{ng2000algorithms}, a 5x5 grid world. The agent is in one of the cells of the grid (whose coordinates is the state) and can choose at each step one of the four compass directions (the action). With probability 0.9, the agent moves in the intended direction. With probability 0.1, the direction is randomly drawn. The reward optimized by the expert is 0 everywhere except in the upper-right cell, where it is 1. For every policy, an episode ends when the upper right cell is attained, or after 20 steps. At the start of each episode, the agent is in the lower-left corner of the grid (the opposite of where the reward is). Both the state and action spaces are finite and of small cardinality. Hence, the chosen feature functions $\phi$ and $\psi$ are the typical features of a tabular representation: 0 everywhere except for the component corresponding to the current state (-action pair).\\

Both \citet{abbeel2004apprenticeship}'s algorithm (from now on referred to as the MC variant) and our adaptation (referred to as the LSTD variant) are tested side by side. The MDP solver of the MC variant is LSPI with a sample source that covers the whole grid (each state has a mean visitation count of more than 150) and draws its action randomly. Both $\mu^{\pi^{(j)}}(s_0)$ and $\mu^{\pi_E}(s_0)$ are computed thanks to a Monte-Carlo estimation with enough samples to make the variance negligible. We consider both these computations as perfect theoretical solvers for all intended purpose on this toy problem. We thus are in the case intended by \citet{abbeel2004apprenticeship}. On the other hand the LSTD variant is used without accessing a simulator. It uses LSPI and LSTD-$\mu$, fed only with the expert's transitions (although we could also use non expert transitions  to compute intermediate policies, if available). This corresponds to a real-life setting where data generation is expensive and the system cannot be controlled by an untrained machine.\\
